---
title: "workflow"
author: "Josh Erickson"
date: "11/8/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(latex2exp)
knitr::opts_chunk$set(dev.args = list(png = list(type = "cairo")),fig.width = 10, fig.height = 5,
                      warning = FALSE, message = FALSE, error = FALSE)
library(patchwork)
library(tidyverse)
library(GGally)
library(ggtext)
library(sasLM)
library(janitor)
library(car)
library(broom)
library(ggpubr)
library(rstatix)
library(readxl)
thematic::thematic_on(font = 'Montserrat')
source('utils.R')
library(sf)
pibo_sites <- read_sf('pibo_sites.gpkg', 'pibo_sites')

gr_95_new <- read_xlsx('R1WestPIBO110521.xlsx') %>% 
  clean_names()

gr_95 <- read_csv('SummaryStatsKendall_gr_95.csv')[,1:36] %>% 
  na.omit() %>%
  clean_names() %>% 
  right_join(gr_95_new %>% select(site_id), by = 'site_id') %>%
  mutate(precip_cut = cut_interval(ave_annual_precip,n=6),
         shed_cut = cut_interval(shed_areakm2, n = 6),
         mgmt = factor(mgmt))
gr_95 <- gr_95 %>% left_join(pibo_sites %>% st_drop_geometry() %>% 
                               select(MAP_UNIT_N, site_id), by = 'site_id')

```

## TL;DR

Below are the results of the ANCOVA model for mean bankfull and mean
bank angle. There were some residuals in mean bankfull that were
affecting the normality of residuals assumption (effects estimating
coeff CIs) so I included a 'filtered' and 'not filtered' analysis for
mean bank full. For mean bank angle everything checked out
`r emo::ji('smile')`. Below is a quick table of the results. I'm sure
we'll be discussing. In my opinion, it's right on the verge (for both
variables) of being an effect; however, it's a smaller effect than what
I expected. What I mean is that precipitation and area dominate the
explained variance in the model (for mean bankfull) so to me the effect
size is smaller than I suspected and changes depending on area and
precip (see estimated marginal mean plots at the very bottom). In short,
there is likely an effect to some degree but I'm not sure that it is
really that meaningful when nested with hydrophysiographic variables
like precip and area (among others). I was expecting something
surprising but it's really quite comparable even when blocking by
ecoregions.

+----------------+------------+------------+------------+------------+
| DV             | IV         | pvalue     | beta       | 95% CI     |
+:==============:+:==========:+:==========:+:==========:+:==========:+
| mean_bf        | mgmt       | 0.043      | beta =     | [0.01,     |
|                |            |            | 0.44       | 0.86]      |
+----------------+------------+------------+------------+------------+
| mean_bf (with  | mgmt       | 0.239      | beta =     | [-0.19,    |
| all residuals) |            |            | 0.28       | 0.74]      |
+----------------+------------+------------+------------+------------+
| mean_bank_an   | mgmt       | 0.116      | beta =     | [-0.65,    |
|                |            |            | 2.64       | 5.93]      |
+----------------+------------+------------+------------+------------+

: Results

**Mean bankfull width**

We fitted a linear model (estimated using OLS) to predict mean_bf with
shed_areakm2, ave_annual_precip and mgmt (formula: mean_bf \~
log(shed_areakm2) + log(ave_annual_precip) + mgmt). The model explains a
statistically significant and substantial proportion of variance (R2 =
0.76, F(3, 298) = 314.23, p \< .001, adj. R2 = 0.76). The model's
intercept, corresponding to shed_areakm2 = 0, ave_annual_precip = 0 and
mgmt = Managed, is at -41.71 (95% CI [-47.78, -35.63], t(298) = -13.52,
p \< .001). Within this model:

*taking three residuals out*

-   The effect of shed_areakm2 [log] is statistically significant and
    positive (beta = 2.86, 95% CI [2.64, 3.08], t(298) = 25.20, p \<
    .001; Std. beta = 1.78, 95% CI [1.64, 1.91])

-   The effect of ave_annual_precip [log] is statistically significant
    and positive (beta = 5.70, 95% CI [4.83, 6.57], t(298) = 12.87, p \<
    .001; Std. beta = 1.13, 95% CI [0.97, 1.30])

-   The effect of mgmt [Reference] is statistically significant and
    positive (beta = 0.44, 95% CI [0.01, 0.86], t(298) = 2.03, p =
    0.043; Std. beta = 0.12, 95% CI [2.58e-03, 0.23])

*including all residuals*

-   The effect of shed_areakm2 [log] is statistically significant and
    positive (beta = 2.99, 95% CI [2.75, 3.24], t(301) = 24.18, p \<
    .001; Std. beta = 1.79, 95% CI [1.66, 1.93])

-   The effect of ave_annual_precip [log] is statistically significant
    and positive (beta = 5.75, 95% CI [4.79, 6.71], t(301) = 11.80, p \<
    .001; Std. beta = 1.08, 95% CI [0.91, 1.25])

-   The effect of mgmt [Reference] is statistically non-significant and
    positive (beta = 0.28, 95% CI [-0.19, 0.74], t(301) = 1.18, p =
    0.239; Std. beta = 0.07, 95% CI [-0.04, 0.19])

**Mean bank angle**

We fitted a linear model (estimated using OLS) to predict mean_bank_an
with shed_areakm2, ave_basin_elev and mgmt (formula: mean_bank_an \~
log(shed_areakm2) + log(ave_basin_elev) + mgmt). The model explains a
statistically significant and substantial proportion of variance (R2 =
0.28, F(3, 301) = 38.80, p \< .001, adj. R2 = 0.27). The model's
intercept, corresponding to shed_areakm2 = 0, ave_basin_elev = 0 and
mgmt = Managed, is at 249.88 (95% CI [191.84, 307.92], t(301) = 8.47, p
\< .001). Within this model:

-   The effect of shed_areakm2 [log] is statistically significant and
    positive (beta = 8.27, 95% CI [6.62, 9.91], t(301) = 9.89, p \<
    .001; Std. beta = 1.11, 95% CI [0.86, 1.35])
-   The effect of ave_basin_elev [log] is statistically significant and
    negative (beta = -23.24, 95% CI [-31.22, -15.26], t(301) = -5.73, p
    \< .001; Std. beta = -0.86, 95% CI [-1.17, -0.54])
-   The effect of mgmt [Reference] is statistically non-significant and
    positive (beta = 2.64, 95% CI [-0.65, 5.93], t(301) = 1.58, p =
    0.116; Std. beta = 0.16, 95% CI [-0.06, 0.38])

## Finding/Exploring/Selecting Covariates

```{r}
gr_95 %>% count(mgmt)

cortest_cols <- function(data, dv, cols){
  
  data_cov <- data %>% select({{cols}})
  data_dv <- data %>% select({{dv}}) %>% pull()
  
  final_data <- data.frame()
  
for(i in 1:length(data_cov)){
  
y <- data_cov[i] %>% pull()
corT <- broom::tidy(cor.test(as.numeric(data_dv), as.numeric(y)))
corT$estimate <- corT$estimate
corT$pvalue <- corT$p.value
corT <- corT %>% mutate(var = names(data_cov[i]))
final_data <- plyr::rbind.fill(final_data, corT)

}

  final_data
}  

mean_bf_cor <- cortest_cols(gr_95, mean_bf, c(drain_density:pct_mtbs_high))

mean_bank_angle_cor <- cortest_cols(gr_95, mean_bank_an, c(drain_density:pct_mtbs_high))

variables_to_check_bf <- mean_bf_cor %>% 
  filter(pvalue <= 0.05)

variables_to_check_angle <- mean_bank_angle_cor %>% 
  filter(pvalue <= 0.05)

#Now visualise and go from there

library(GGally)

ggpairs(gr_95 %>% select(mean_bank_an, variables_to_check_angle$var, mgmt),
        mapping = aes(color = mgmt), upper = list(continuous = wrap("cor", size = 3)),
        lower = list(continuous = wrap("smooth", method = "lm", se = F,
                                               alpha = 0.6)))+
  custom_theme() + labs(title = 'Comparing Mean Bank Angle with Potential Covariates')

ggpairs(gr_95 %>% select(mean_bf, variables_to_check_bf$var[6:10], mgmt),
        mapping = aes(color = mgmt), upper = list(continuous = wrap("cor", size = 3)),
        lower = list(continuous = wrap("smooth", method = "lm", se = F,
                                               alpha = 0.6)))+
  custom_theme() + labs(title = 'Comparing Mean Bankfull Width with Potential Covariates')

ggpairs(gr_95 %>% select(mean_bf, variables_to_check_bf$var[1:5], mgmt),
        mapping = aes(color = mgmt), upper = list(continuous = wrap("cor", size = 3)),
        lower = list(continuous = wrap("smooth", method = "lm", se = F,
                                               alpha = 0.6)))+
  custom_theme() + labs(title = 'Comparing Mean Bankfull Width with Potential Covariates')


```

Looks like `shed_areakm2` and `ave_annual_precip` for `mean_bf` and
`shed_areakm2` and `ave_basin_elev` for `mean_bank_an`. These have
decent $r$ values and also make sense causally so we'll add them to the
model.

## Check Assumptions

Since we are using linear regression (LR) we need to make sure our
models follow LR assumptions; must be linear, multicollinearity, iid,
homoscedastic and normality of residuals. In addition, we need to check
assumptions for ANCOVA; homogeneity of slopes and independence between
iv and covariate(s). We'll start with linearity followed by
multicollinearity and then run the models so we can finish the tests.

### Linearity

We'll take a look at the dv's and the associated covariates to see if
the are linear.

```{r}
#mean bf
gr_95 %>%
  ggplot(aes(shed_areakm2, mean_bf)) +
  geom_point() + 
  geom_smooth(method = 'lm') + custom_theme()
  
#looks a little curves towards the lower watershed areas; log transform

gr_95 %>% 
  ggplot(aes(shed_areakm2, mean_bf)) +
  geom_point() + 
  scale_x_log10() +
  geom_smooth(method = 'lm') + custom_theme()

# precip 

gr_95 %>% 
  ggplot(aes(ave_annual_precip, mean_bf)) +
  geom_point() + 
  scale_x_log10() +
  geom_smooth(method = 'lm') + custom_theme()

#much better

#now to mean bank angle

gr_95 %>%
  ggplot(aes(shed_areakm2, mean_bank_an)) +
  geom_point() + 
  geom_smooth(method = 'lm') + custom_theme()

#sure...

gr_95 %>% filter(shed_areakm2 >= 2) %>%
  ggplot(aes(shed_areakm2, mean_bank_an, color = mgmt)) +
  geom_point() + 
  scale_x_log10()+
  geom_smooth(method = 'lm') + custom_theme()

# much better

```

## Multicollinearity

We'll use the vif() (variance inflation factor) to check and make sure
we don't have any multicollinearity.

```{r}

#create our model for mean_bf
#anything over five is a problem
mod_bf <- lm(mean_bf~log(shed_areakm2)+log(ave_annual_precip)+mgmt, data = gr_95)
car::vif(mod_bf)

#looks good

#create our model for mean_bank_angle
#anything over five is a problem
mod_bank <- lm(mean_bank_an~log(shed_areakm2)+log(ave_basin_elev)+mgmt, data = gr_95)
car::vif(mod_bank)

#looks good

```

## iid

Here we'll lump iid and independence between iv and covariate(s). We are
assuming iid since all the observations are independent of each other
and are far enough away; identically distributed is mostly followed
since the observations are 'response' reaches (less than 4% gradient)
but may have different hydrological regimes, i.e. rain-on-snow vs
snowmelt. The iv is not dependent on the covariates since the covariates
are not a function of the iv, i.e. precipitation doesn't care if it's a
managed area or reference; it just happens.

## Models

We'll run the models for ANCOVA below for both dv's.

```{r}

#for mean_bf
mod_bf <- lm(log(mean_bf)~log(shed_areakm2)+log(ave_annual_precip)+mgmt, data = gr_95)
Anova(mod_bf, type = 'III')


#for mean_bank_an
mod_bank <- lm(mean_bank_an~log(shed_areakm2)+log(ave_basin_elev)+mgmt, data = gr_95)
Anova(mod_bank, type = 'III')


```

## Residual assumptions

Now that we have the models we can check the residual assumptions;
homogeneity of variance and normally distributed.

```{r}
library(performance)

#mean_bf
check_normality(mod_bf)
#not normal

check_heteroscedasticity(mod_bf)
#not hetereoscedastic 

#mean_bank_an

check_normality(mod_bank)
#normal

check_heteroscedasticity(mod_bank)
#homoscedastic 
```

Let's look at the `mod_bf` model and see what's going on.

```{r}
qqPlot(mod_bf)
ggplot() +
  geom_density(aes(mod_bf$residuals[-75]))

```

As you can see we have some high positive residuals that are long in
tails.

Let's look at the spatial part of the residuals and see if there is
something going on.

```{r}
pibo_sites$residuals <- mod_bf$residuals

mt <- AOI::aoi_get(state = "MT")
id <- AOI::aoi_get(state = "ID")
mt_id <- bind_rows(mt, id)

ggplot() + 
  geom_sf(data = mt_id, fill = NA, alpha = 0) + 
  geom_sf(data = pibo_sites, aes(color = residuals), size = 2.75)+
  scale_color_gradientn(colors = wesanderson::wes_palette('Zissou1')) +
  theme_void(base_size = 16) + coord_sf(xlim = c(-117,-112), ylim = c(49.1, 45)) +labs(color = 'Residuals',                                                                       title = 'Maps showing model residuals for Y = 	&beta;<sub>0</sub> + 	&beta;<sub>1</sub>*Precipitation (mm) + &beta;<sub>2</sub>*Watershed Area (km<sup>2</sup>) + &beta;<sub>3</sub>*Group<br>') +facet_wrap(~mgmt) +
  theme(plot.title = element_markdown(size = 12))


```

I think just taking out the location `r gr_95[75,]$stream`.

```{r}
#now let's do the qqPlot() without these

mod_bf_filt <- lm(log(mean_bf)~log(shed_areakm2)+log(ave_annual_precip)+mgmt, data = gr_95)
gr_95$residuals <- mod_bf_filt$residuals
mod_bf_filt <- lm(log(mean_bf)~log(shed_areakm2)+log(ave_annual_precip)+mgmt, data = gr_95[-c(75),])
qqPlot(mod_bf_filt)
check_normality(mod_bf_filt)
```

Now check the assumptions of the model.

```{r}
check_normality(mod_bf_filt)
#normal

check_heteroscedasticity(mod_bf_filt)
#homoscedastic
```

Better. But should we do this... Well let's see what happens.

Now we need to see if the regression lines are homogeneous.

## Homogeneity of regression slopes assumption

If the covariate and iv interaction are significant then we have
violated the assumption.

```{r}
mod_bf_precip <- lm(log(mean_bf)~mgmt*log(ave_annual_precip), data = gr_95)
summary(mod_bf_precip)
#passed

mod_bf_shed <- lm(log(mean_bf)~mgmt*log(shed_areakm2), data = gr_95)
summary(mod_bf_shed)
#passed 


mod_ba_elev <- lm(mean_bank_an~mgmt*log(ave_basin_elev), data = gr_95)
summary(mod_bf_elev)
#passed

```

## Run the final models

```{r}
#for mean_bf filter out high residuals
mod_bf_filt <- lm(mean_bf~log(shed_areakm2)+log(ave_annual_precip)+MAP_UNIT_N+mgmt, data =gr_95[-75,])
Anova(mod_bf_filt, type = 'III')

#for mean_bf WITHOUT filtering high residuals
mod_bf <- lm(mean_bf~log(shed_areakm2)+log(ave_annual_precip)+MAP_UNIT_N+mgmt, data = gr_95)
Anova(mod_bf, type = 'III')
#for mean bank angle
mod_bank <- lm(mean_bank_an~log(shed_areakm2)+log(ave_basin_elev)+MAP_UNIT_N+mgmt, data = gr_95)
summary(mod_bank)
Anova(mod_bank, type = 'III')

```

## Now we'll look at estimated marginal means

We now need to look at the estimated marginal means (emm) along with the
confidence intervals to see how the groups compare.

```{r}
library(emmeans)
library(car)

bf.rg <- ref_grid(mod_bf_filt, at = list(shed_areakm2 = seq(1, 100, 5)))

emm <- plot(bf.rg, by = 'mgmt') + custom_theme()

emm_gg <- as.data.frame(emm$data) %>% tibble()


p1 <- emm_gg %>% mutate(shed_area = ifelse(mgmt %in% 'Reference', shed_areakm2+1.25, shed_areakm2-1.25),
                        the.emmean = exp(the.emmean),
                        lower.CL = exp(lower.CL),
                        upper.CL = exp(upper.CL)) %>% filter(the.emmean >0) %>% 
  ggplot() +
  #geom_point(data = gr_95 , aes(mean_bf, shed_areakm2, color = mgmt), size = 3, alpha = 0.2)+
  geom_errorbar(aes(the.emmean, shed_areakm2, color=mgmt,xmin = lower.CL, xmax = upper.CL),
  position = position_dodge2(width = 0.005, padding = 0.005), size = .75) + 
  custom_theme(font_size = 10) + labs(x = 'Predicted Mean Bankfull Width (m)', 
                        y = 'Watershed Area km2',
                        color = "Group",
                        title = 'Comparing estimated means by Watershed Area (km<sup>2</sup>) (without high residuals)',
                        subtitle = 'Average annual precipitation is held constant at 1193 (mm)') +
  geom_point(aes(the.emmean, shed_area), size = 1)  +
  geom_line(aes(the.emmean, shed_areakm2,color = mgmt, group = mgmt),
  position = position_dodge2(width = 0.1, padding = 0), alpha = 0.25)
p1 
managed <- emm_gg[1:20,]
ref <- emm_gg[21:40,]

managed <- managed %>% mutate(refCL = ref$lower.CL, 
                              highCL = ref$upper.CL,
                              emmean = ref$the.emmean,
                              se_emmean = ref$SE)
p2 <- managed %>% mutate(cross_zero = refCL-the.emmean,
                         model = 'Filtered\npvalue = 0.043') %>% 
  ggplot(aes(shed_areakm2, cross_zero)) +
  geom_col() + 
  labs(x = 'Watershed Area km<sup>2</sup>',
       y = 'difference',
       title = 'Difference between Reference Lower CI and Estimated Mean of Managed (without high residuals)') +
  custom_theme(font_size = 10)

p2

 bf.rg <- ref_grid(mod_bf, at = list(shed_areakm2 = seq(1, 100, 5)))

emm <- plot(bf.rg, by = 'mgmt') + custom_theme()

emm_gg <- as.data.frame(emm$data) %>% tibble()


p3 <- emm_gg %>% mutate(shed_area = ifelse(mgmt %in% 'Reference', shed_areakm2+1.25, shed_areakm2-1.25)) %>% filter(the.emmean >0) %>% 
  ggplot() +
  #geom_point(data = gr_95 , aes(mean_bf, shed_areakm2, color = mgmt), size = 3, alpha = 0.2)+
  geom_errorbar(aes(the.emmean, shed_areakm2, color=mgmt,xmin = lower.CL, xmax = upper.CL),
  position = position_dodge2(width = 0.005, padding = 0.005), size = .75) + 
  custom_theme(font_size = 10) + labs(x = 'Predicted Mean Bankfull Width (m)', 
                        y = 'Watershed Area km2',
                        color = "Group",
                        title = 'Comparing estimated means by Watershed Area (km<sup>2</sup>) (with high residuals)',
                        subtitle = 'Average annual precipitation is held constant at 1193 (mm)') +
  geom_point(aes(the.emmean, shed_area), size = 1) 
p3 


```


## Linear mixed models  

```{r}

library(lme4)
library(resourceviz)
cairo_view()
ggplot(gr_95_ts) + 
  geom_point(aes(yr, bf, group = site_id)) + 
  stat_smooth(geom = 'line',aes(yr,bf, group = site_id, color = mgmt), method = 'lm', se = F, alpha = 0.5, size = 1) +
  geom_smooth(aes(yr, bf), col = 'red', method = 'lm', size = 1.5) +
  #scale_color_gradientn(colors = hcl.colors('Zissou1', n = 11)) +
  facet_wrap(MAP_UNIT_N~mgmt) +
  custom_theme() + 
  labs(x = 'Sample Year', y = 'Bankfull Width (m)')


ggplot(gr_95_ts) + 
  geom_point(aes(yr, bf, group = site_id), alpha = 0.05) + 
  stat_smooth(geom = 'line',aes(yr,bf,color = MAP_UNIT_N), method = 'lm', se = F, size = .5) +
  geom_smooth(aes(yr, bf), col = 'black', method = 'lm', size = .5, linetype = 2) +
  #scale_color_gradientn(colors = hcl.colors('Zissou1', n = 11)) +
  custom_theme() + 
  facet_wrap(~mgmt) +
  labs(x = 'Sample Year', y = 'Bankfull Width (m)')



library(jtools)
lmm_bf <- lmer(bf ~ log(shed_areakm2) + log(ave_annual_precip) + mgmt + (1 | MAP_UNIT_N/site_name), gr_95_ts, REML = F)
summ(lmm_bf)


lmm_ba <- lmer(bank_angle ~ log(shed_areakm2) + log(ave_basin_elev) + mgmt + (1 | MAP_UNIT_N/site_name),gr_95_ts, REML = F)
summ(lmm_ba)
rsb <- list(psi2propII(smoothPsi), psi2propII(smoothPsi, k = 2.28))
rfm3 <- update(lmm_bf, rho.sigma.b = rsb)

plot(rfm3)

compare(lmm_bf, rfm3, show.rho.functions = F)
library(robustlmm)
lmm_bf <- rlmer(bf ~ log(shed_areakm2) + log(ave_annual_precip) + mgmt + (1 | MAP_UNIT_N/site_name), gr_95_ts, REML = F)

lmm_bf2 <- lmer(bf ~ log(shed_areakm2) + log(ave_annual_precip) + mgmt + (1 | MAP_UNIT_N/site_name), gr_95_ts, REML = F)

summ(lmm_bf)
summary(lmm_bf)
anova(lmm_bf)
coefs <- data.frame(coef(summary(lmm_bf2)))

# get coefficients from robust model to extract t-values
coefs.robust <- coef(summary(lmm_bf))
rownames(coefs.robust) <- NULL
# calculate p-values based on robust t-values and non-robust approx. DFs
p.values <- 2*pt(abs(coefs.robust[,3]), abs(coefs$t.value), lower=FALSE)


lmm_ba <- rlmer(bank_angle ~ log(shed_areakm2) + log(ave_basin_elev) + mgmt + (1 | MAP_UNIT_N/site_name),gr_95_ts, REML = F)
summ(lmm_ba)


ggplot(gr_95_ts) + 
  geom_point(aes(yr, bank_angle, group = site_id), alpha = 0.05) + 
  stat_smooth(geom = 'line',aes(yr,bank_angle,color = MAP_UNIT_N), method = 'lm', se = F, size = .5) +
  geom_smooth(aes(yr, bank_angle), col = 'black', method = 'lm', size = .5, linetype = 2) +
  #scale_color_gradientn(colors = hcl.colors('Zissou1', n = 11)) +
  custom_theme() + 
  facet_wrap(~mgmt) +
  labs(x = 'Sample Year', y = 'Bankfull Width (m)')

ggplot(gr_95_ts %>% filter(mgmt == 'Managed')) + 
  geom_point(aes(yr, bf, group = site_id), alpha = 0.05) + 
  stat_smooth(geom = 'line',aes(yr,bf,color = harv_group), method = 'lm', se = F, size = .5) +
  scale_color_manual(values = c('blue', 'green', 'red')) +
  custom_theme() +
  facet_wrap(~MAP_UNIT_N)


library(performance)
check_normality(lmm_bf2)
check_heteroscedasticity(lmm_bf)
check_normality(lmm_ba)
check_heteroscedasticity(lmm_ba)
plot(lmm_ba)
plot(lmm_bf)
qqPlot(lmm_bf)
plot(lmm_bf)
plot(residuals(lmm_bf, 'weighted'))

shapiro.test(residuals(lmm_bf))
shapiro.test(rnorm(1000))
```



```{r}
gr_95_ts <- gr_95_all %>%
  right_join(gr_95 %>% select(site_id, drain_density:pct_nfs, mean_bf:shed_cut,MAP_UNIT_N, -mgmt)) %>%
  mutate(impact_fire = pct_mtbs_mod+pct_mtbs_high,
         impact_harvest = pct_high_harv_int+pct_mod_harv_int,
         harv_group = ifelse(impact_harvest <= 5, 'Low', ifelse(impact_harvest > 5 & impact_harvest <= 15, 'Mod', 'High')),
         harv_group = factor(harv_group, levels = c('Low', 'Mod', 'High')),
         fire_group = ifelse(impact_fire <= 5, 'Low', ifelse(impact_fire > 5 & impact_fire <= 18, 'Mod', 'High')),
         fire_group = factor(fire_group, levels = c('Low', 'Mod', 'High')))

gr_95_ts <- gr_95_ts %>%
  group_by(site_id) %>% summarise(sd_bf = sd(bf, na.rm = T),
                                  meanbf = mean(bf, na.rm = T),
                                  coef_bf = sd_bf/meanbf,
                                  coef_bfl = sqrt(exp(sd(log(bf), na.rm = TRUE)^2)-1),
                                  sd_ba = sd(bank_angle, na.rm = T),
                                  meanba = mean(bank_angle, na.rm = T),
                                  coef_ba = sd_ba/meanba,
                                  coef_bal = sqrt(exp(sd(log(bank_angle), na.rm = TRUE)^2)-1)) %>%
  ungroup() %>%
  left_join(gr_95_ts, by = 'site_id') 

gr_95 <- gr_95_ts %>% group_by(site_id) %>% slice(n = 1) %>% ungroup()

lmm_bf <- lmer(bf ~ log(shed_areakm2) + log(ave_annual_precip) + harv_group + (1 | MAP_UNIT_N/site_name), gr_95_ts %>% filter(mgmt == 'Managed'), REML = F)
summ(lmm_bf)


lmm_ba <- lmer(bank_angle ~ log(shed_areakm2) + log(ave_basin_elev) + harv_group + (1 | MAP_UNIT_N/site_name),gr_95_ts %>% filter(mgmt == 'Managed'), REML = F)
summ(lmm_ba)


lmm_bf <- lmer(bf ~ log(shed_areakm2) + log(ave_annual_precip) + fire_group + (1 | MAP_UNIT_N/site_name), gr_95_ts %>% filter(mgmt == 'Reference'), REML = F)
summ(lmm_bf)


lmm_ba <- lmer(bank_angle ~ log(shed_areakm2) + log(ave_basin_elev) + fire_group + (1 | MAP_UNIT_N/site_name),gr_95_ts %>% filter(mgmt == 'Reference'), REML = F)
summ(lmm_ba)
```



























